On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2022-10-24 manuscript/manuscript.tex
smoothed section 2

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample of size~$n$ from a continuous
distribution. The null hypothesis $H_0$ is that $X_i$'s follow distribution~$F$.
Let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function. The KS test statistic is
\begin{equation}
  \label{eq:ks_standard}
  D_n = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The asymptotic distribution of $D_n$ under $H_0$ is independent of the
distribution $F$. As $n \to \infty$, $\sqrt{n} D_n$ converges in distribution to
the supremum of standard Brownian bridge \citep{kolmogorov1933sulla}. For large
samples, the tests can be performed with a table \citep{smirnov1948table}.
Critical values for small samples ($n \le 35$) has also been given
\citep{Massey}. The KS test is available in popular
statistical software packages, such as function \texttt{ks.test} in R package
\textsf{stats} \citep{R, Marsaglia}.

@@

The null distribution of the KS statistic changes when the hypothesized
distribution contains fitted parameters.
\eds{what is the null distribution when not using fitted parameters}.
In this scenario, the null hypothesis
is $H_0$: the random sample $X_1, \ldots, X_n$ comes from a continuous
distribution $F_{\theta}$ with unspecified parameter vector $\theta$.
Let $\hat\theta_n$ be an estimator of $\theta$, which could be, for example, a
maximum likelihood estimator or moment estimator. The test statistic is
\begin{equation}
  \label{eq:ks_fitted}
  D = \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\end{equation}
Since $F_{\hat\theta_n}$ is not the same as the true data generating
$F_\theta$, the null distribution of $D$ obtained in existing implementations in
software packages which assumes completely known $F_\theta$ no longer applies
\citep{Steinskog}.

To demonstrate the consequences of this problem, a simulation is performed. A
random sample $X_1, \ldots, X_n$ is generated from the standard normal
distribution with sample size $n = 100$. The p-values of $1000$ replicate tests
 of normality
are displayed in the Naive plot of Figure~\ref{fig:hist_fitted}.
\eds{how are these p-values calculated?}
Each KS test
was performed using fitted parameters, i.e., the hypothesized distribution is
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance. Since the data is truly generated from a standard normal distribution,
 %with
%seemingly all assumptions met,
a uniform distribution of $U(0, 1)$ would naively be expected for the p-values.
However this should only be expected when the KS test assumptions hold, and they
 do not due to using fitted parameters. Therefore,
there is notable deviation from the uniform distribution.

To fix the problem, parametric bootstrap can be used to approximate the null
distribution of the testing statistic.
\begin{enumerate}
  \item
    Draw a random sample $X_1^*,...,X_n^*$ from the fitted distribution
    $F_{\hat\theta_n}$
  \item
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function $F_n^*$ of $X_1^*, \ldots,
    X_n^*$.
  \item
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) |.
    \]
  \item
    Repeat the previous steps a large number $B$ times and use the empirical
    distribution of $D^*$ to approximate the null distribution of the observed
    statistic.
\end{enumerate}

Nonparametric bootstrap can also be used to approximate the null distribution
of the testing statistic. The procedure is similar, however the resampling is
performed with the empirical distribution instead of the fitted parametric
distribution and there is a correction for bias that is required
\citep{Babu}.
\begin{enumerate}
  \item
    Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
    with replacement
  \item
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
    \]
    where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
    bias term \citep{Babu}
  \item
    Repeat the previous steps a large number $B$ times and use the empirical
    distribution of $D^*$ to approximate the null distribution of the observed
    statistic.
\end{enumerate}
The p-value can be calculated by counting the number of bootstrap KS
statistics greater than or equal to the observed KS statistic, and then dividing
by the number of bootstrap samples. Figure~\ref{fig:hist_fitted} displays the
results of from our simulations. We would expect the distribution of p-values
to be uniform in the case where the KS test holds its size. It is clear from the
figure that both parametric and nonparametric bootstrap processes correct for
the problem of fitted parameters. The plots for the bootstrapped p-values appear
to be $U(0,1)$, unlike the naive p-values.

@@

>>>>

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample of size~$n$ from a continuous
distribution. The null hypothesis $H_0$ is that $X_i$'s follow distribution~$F$.
Let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function. The KS test statistic is
\begin{equation}
  \label{eq:ks_standard}
  D_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The asymptotic distribution of $D_n$ under $H_0$ is independent of the
distribution~$F$. As $n \to \infty$, $D_n$ converges in distribution to
the supremum of standard Brownian bridge \citep{kolmogorov1933sulla}. For large
samples, the tests can be performed with a table \citep{smirnov1948table}.
Critical values for small samples ($n \le 35$) has also been given
\citep{Massey}. The KS test is available in popular
statistical software packages, such as function \texttt{ks.test} in R package
\textsf{stats} \citep{R, Marsaglia}.

@@

The null hypothesis of a goodness-of-fit test is often a composite hypothesis
instead of a single hypothesis. That is, the hypothesized distribution is a
family of distributions with unspecified parameters instead of a specific member
in this family. Let $F_\theta$ be a family of distributions indexed by parameter
vector~$\theta$. The null hypothesis is
\[
  H_0: \text{the random sample $X_1, \ldots, X_n$
    comes from a distribution $F_\theta$ with some $\theta$.}
\]
Since $\theta$ is unknown, one would naturally estimate $\theta$ by
$\hat\theta_n$ from, for example, maximum likelihood or methods of moments. The
KS statistic would then be computed as
\begin{equation}
  \label{eq:ks_fitted}
  D_n = \sqrt{n} \sup_x | F_n(x) - F_{\hat\theta_n}(x) |.
\end{equation}
An overly large $D_n$ still indicates evidence to reject~$H_0$. To get the
p-value of the observed $D_n$, however, the null distribution is not the same as the
of $D_n$, however, is no longer the same as that in the standard
case~\eqref{eq:ks_standard}. If the same null distribution were used, one would
be testing a different $H_0$ that the random sample comes from the specific
member distribution $F_{\hat\theta}$.

The consequence of using the wrong null distribution for the KS statistic $D_n$
in~\eqref{eq:ks_fitted} can be illustrated through a simple simulation study. A
random sample $X_1, \ldots, X_n$ was generated from the standard normal
distribution with sample size $n = 100$. The sample mean and sample standard
deviation were used as $\hat\theta_n$. The p-value of the testing
statistic~$D_n$ in~\eqref{eq:ks_fitted} was obtained by naively calling the
\texttt{ks.test} function in~R with the fitted normal distribution as the
hypothesized distribution. That is, the hypothesized distribution was
$N(\bar X, s^2)$ where $\bar X$ is the sample mean and $s^2$ is the sample
variance. This experiment was repeated for 1000 times and the histogram of the
1000 p-values are displayed in Figure~\ref{fig:hist_fitted}. If the test were
valid, the p-values would be uniformly distributed in the unit
interval. Clearly, the pattern of the 1000 p-values are very different from what
one would expect to see from a 1000 draws from the standard uniform
distribution.

To fix the problem, parametric bootstrap can be used to approximate the null
distribution of the testing statistic~$D_n$. For a large number~$B$, the
parametric bootstrap procedure is as follows.
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the fitted distribution
  $F_{\hat\theta_n}$
\item
  Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
\item
  Obtain the empirical distribution function $F_n^*$ of
  $X_1^*, \ldots,  X_n^*$.
\item
  Calculate bootstrap KS statistic
  \[
    D_n^* = \sqrt{n} \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) |.
  \]
\item
  Repeat the previous steps $B$ times and use the empirical
  distribution of $D^*$ to approximate the null distribution of the observed
  statistic.
\end{enumerate}
The p-value of $D_n$ is approximated by the proportion of $D_n^*$'s that are
greater than or equal to~$D_n$.

Nonparametric bootstrap can also be used to approximate the null distribution
of the testing statistic with the bias correction from \citet{Babu}. The
procedure is as follows.
\begin{enumerate}
\item
  Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
  with replacement
\item
  Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
\item
  Obtain the empirical distribution function of the random sample $F_n^*$
\item
  Calculate bootstrap KS statistic
  \[
    D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
  \]
  where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known
  bias term \citep{Babu}
\item
  Repeat the previous steps a large number $B$ times and use the empirical
  distribution of $D^*$ to approximate the null distribution of the observed
  statistic.
\end{enumerate}
The p-value of $D_n$ is, again, approximated by the proportion of $D_n^*$'s that
are greater than or equal to~$D_n$.

@@

In the same simulation study, p-values were obtained for the 1000 replicates
with both the parametric and nonparametric bootstrap procedures. The center and
right panels of Figure~\ref{fig:hist_fitted} display the histograms of the 1000
p-values from the parametric and nonparametric bootstrap procedures,
respectively. This time, both samples of the p-values appear to come from the
standard uniform distributions.
